### 2015-11-30 ch02-qlogms/excitation-design.tex
Some info on periodicity
Example: show period of quantized GMS

  Consequently, the complete multisine has a period $T_{u(t)}$ that is the common multiple of the period of all its components:
  \begin{equation}
    T_{u(t)}  = \lcm \set{T_1, \ldots, T_F}
      = \lcm \set{f_1^{-1}, \ldots, f_F^{-1}}
    \text{.}
  \end{equation}
  \begin{definition}[Least common (integer) multiple] \label{def:lcm-real}
  The least common multiple of the set of values $T_k \in \RealNumbers$ for $k \in \set{1, \ldots, F}$ is  
  \begin{equation}
    \lcm \set{T_1, \ldots, T_F} \isdef
      \min \Set{T |\forall k \in \set{1,\ldots,F}, \exists n_k \in \NaturalNumbers: n_k T_k = T }
   \end{equation}
   if such a minimum exists.
   \end{definition}
  \begin{example}
   Consider a generalized multisine where $f_1 = 1 \unit{Hz}$ and $f_2 = \tfrac{2}{3} \unit{Hz}$.
   The corresponding periods are $T_1 = f_1^{-1} = 1 \unit{s}$ and $T_2 = f_2^{-1} = 1.5 \unit{s}$.
   The overall period of the multisine is $T = n_1 T_1 = n_2 T_2 = 3 \unit{s}$, i.e. $n_1 = 3$ and $n_2=2$.
  \end{example}

  \begin{example}
   Consider the generalized multisine of the previous example to which a spectral component is added at $f_3 = \pi \unit{Hz}$, or equivalently one with a period of $T_3 = \pi^{-1} \unit{s}$.
   Then it is not possible to determine integer values $n_1$, $n_2$, $n_3$ such that $n_1 T_1 = n_2 T_2 = n_3 T_3$ since $T_3$ is an irrational number of seconds (unlike $T_1$ and $T_2$).
   Consequently, such a generalized `multisine' is not periodic.
  \end{example}

>>>>

  Consequently, the complete multisine has a period $T_{u(t)}$ that is the common multiple of the period of all its components:
  \begin{equation}
    T_{u(t)}  = \lcm \set{T_1, \ldots, T_F}
      = \lcm \set{f_1^{-1}, \ldots, f_F^{-1}}
    \text{.}
  \end{equation}
  \begin{definition}[Least common (integer) multiple] \label{def:lcm-real}
  The least common multiple of the set of values $T_k \in \mathbb{F}\without\set{0}$ for $k \in \mathbb{K} \isdef \set{1, \ldots, F}$ is  
  \begin{equation}
    \lcm \set{T_1, \ldots, T_F} \isdef
      \min \Set{T |\forall k \in \mathbb{K}, \exists n_k \in \NaturalNumbersWithoutZero: n_k T_k = T }
   \end{equation}
   if such a minimum exists.
   When $\mathbb{F} \in \Set{\NaturalNumbers, \Integers, \RationalNumbers}$, a minimum always exists.
   For $\mathbb{F} = \RealNumbers$, a minimum is not guaranteed to exist (see below).
   % more info: http://mathworld.wolfram.com/LeastCommonMultiple.html
   \end{definition}
  \begin{example}
   Consider a generalized multisine where $f_1 = 1 \unit{Hz}$ and $f_2 = \tfrac{2}{3} \unit{Hz}$.
   The corresponding periods are $T_1 = f_1^{-1} = 1 \unit{s}$ and $T_2 = f_2^{-1} = 1.5 \unit{s}$.
   The overall period of the multisine is $T = n_1 T_1 = n_2 T_2 = 3 \unit{s}$, i.e. $n_1 = 3$ and $n_2=2$.
  \end{example}

  \begin{example}\label{eg:excitation:non-periodic}
   Consider the generalized multisine of the previous example to which a spectral component is added at $f_3 = \pi \unit{Hz}$, or equivalently one with a period of $T_3 = \pi^{-1} \unit{s}$.
   Then it is not possible to determine integer values $n_1$, $n_2$, $n_3$ such that $n_1 T_1 = n_2 T_2 = n_3 T_3$ since $T_3$ is an irrational number of seconds (unlike $T_1$ and $T_2$).
   Consequently, such a generalized `multisine' is not periodic.
  \end{example}
  \begin{remark}
  In practical applications each frequency will be represented as either a fixed-point or floating point number with a limited numerical precision $\delta$.
  Hence, each frequency line $f_k$ will be approximated by a line at $\round{f_k \delta^{-1}} \delta \isdef m_k \delta$ where $m_k \in \Integers$.
  The resulting multisine has a period 
  \begin{equation}
     T 
     = \lcm \Set{T_k |  k \in \mathbb{K}} 
     = \lcm \Set{m_k^{-1} \delta^{-1} |  k \in \mathbb{K}}
     = \delta^{-1} \lcm \Set{m_k^{-1} |  k \in \mathbb{K}}
     \text{.}
  \end{equation}
  The latter $\lcm$ always exists because $m_k^{-1}  \in \RationalNumbers$ as $m_k \in \Integers$.
  The resulting period $T$ might, however, might be too long for practical use.
  E.g. when $\delta \approx 2.2\cdot 10^{-16}$ (\code{double} precision), the period of a multisine with components as in Example~\ref{eg:excitation:non-periodic} is $T \approx 2.8 \cdot 10^{14} \unit{s} \approx 9 \cdot 10^{6} \unit{year}$.
  \end{remark}

### 2015-11-30 config/glossaries.inc.tex
Some info on periodicity
Example: show period of quantized GMS

\newsymbolglos{A-SetC}{\ComplexNumbers}{Set of the complex numbers}
\newsymbolglos{A-SetN}{\NaturalNumbers}{Set of the natural numbers}
\newsymbolglos{A-SetRpos}{\PositiveRealNumbers}{Set of the positive real numbers}
\newsymbolglos{A-SetRnozero}{\RealNumbersWithoutZero}{Set of the positive real numbers without the number $0$, i.e. $\RealNumbers \without \set{0}$}
\newsymbolglos{A-SetR}{\RealNumbers}{Set of the real numbers}
\newsymbolglos{A-SetZ}{\IntegerNumbers}{Set of integer numbers}

@@

\newsymbolglos{E-approx}{\approx}{Is approximately equal to \ldots}
\newsymbolglos{E-bydef}{\isdef}{Is defined as \ldots}
\newsymbolglos{E-propto}{\propto}{Is proportional to \ldots}
\newsymbolglos{E-distributed}{\distributedAs}{Is distributed as \ldots}

@@

\newsymbolglos{operator-cardinality}{\numberOf \placeholder}{Number of elements (cardinality) in the set \placeholder}
\newsymbolglos{operator-exp}{\expval{\placeholder}}{Expected value of \placeholder}
\newsymbolglos{operator-ln}{\ln \placeholder}{Natural logarithm of \placeholder}
\newsymbolglos{operator-conj}{\conj \placeholder}{Complex conjugate of \placeholder}
\newsymbolglos{operator-log}{\log_{B} \placeholder}{Logarithm in base $B$ of \placeholder{}  (when B is omitted, the decimal logarithm is implied)}
\newsymbolglos{operator-matrix-ht}{\placeholder^{\HT}}{Hermitian (complex conjugated) transpose of the matrix \placeholder}
\newsymbolglos{operator-matrix-pinv}{\pinv{\placeholder}}{Moore-Penrose pseudo-inverse of the matrix \placeholder}
\newsymbolglos{operator-matrix-tran}{\placeholder^{\TT}}{Transpose of the matrix \placeholder}
\newsymbolglos{operator-round-ceil}{\ceil{\placeholder}}{\placeholder{} rounded upwards (\code{ceil})}
\newsymbolglos{operator-round-floor}{\floor{\placeholder}}{\placeholder{} rounded downwards (\code{floor})}
\newsymbolglos{operator-round-round}{\round{\placeholder}}{\placeholder{} rounded towards the nearest integer (\code{round})}

>>>>

\newsymbolglos{A-SetC}{\ComplexNumbers}{Set of the complex numbers}
\newsymbolglos{A-SetN}{\NaturalNumbers}{Set of the natural numbers}
\newsymbolglos{A-SetNnozero}{\NaturalNumbersWithoutZero}{Set of the natural numbers without the number $0$}
\newsymbolglos{A-SetRpos}{\PositiveRealNumbers}{Set of the positive real numbers}
\newsymbolglos{A-SetRneg}{\NegativeRealNumbers}{Set of negative real numbers}
\newsymbolglos{A-SetRnozero}{\RealNumbersWithoutZero}{Set of the positive real numbers without the number $0$, i.e. $\RealNumbers \without \set{0}$}
\newsymbolglos{A-SetQ}{\RationalNumbers}{Set of the rational numbers}
\newsymbolglos{A-SetR}{\RealNumbers}{Set of the real numbers}
\newsymbolglos{A-SetZ}{\IntegerNumbers}{Set of integer numbers}

@@

\newsymbolglos{E-approx}{\approx}{is approximately equal to \ldots}
\newsymbolglos{E-bydef}{\isdef}{is defined as \ldots}
\newsymbolglos{E-propto}{\propto}{is proportional to \ldots}
\newsymbolglos{E-distributed}{\distributedAs}{is distributed as \ldots}

@@

\newsymbolglos{operator-cardinality}{\numberOf \placeholder}{Number of elements (cardinality) in the set \placeholder}
\newsymbolglos{operator-exp}{\expval{\placeholder}}{Expected value of \placeholder}
\newsymbolglos{operator-ln}{\ln \placeholder}{Natural logarithm of \placeholder}
\newsymbolglos{operator-conj}{\conj \placeholder}{Complex conjugate of \placeholder}
\newsymbolglos{operator-log}{\log_{B} \placeholder}{Logarithm in base $B$ of \placeholder{}  (when $B$ is omitted, $B=10$ is implied)}
\newsymbolglos{operator-matrix-ht}{\placeholder^{\HT}}{Hermitian (complex conjugated) transpose of the matrix \placeholder}
\newsymbolglos{operator-matrix-pinv}{\pinv{\placeholder}}{Moore-Penrose pseudo-inverse of the matrix \placeholder}
\newsymbolglos{operator-matrix-tran}{\placeholder^{\TT}}{Transpose of the matrix \placeholder}
\newsymbolglos{operator-round-ceil}{\ceil{\placeholder}}{\placeholder{} rounded upwards (\code{ceil})}
\newsymbolglos{operator-round-floor}{\floor{\placeholder}}{\placeholder{} rounded downwards (\code{floor})}
\newsymbolglos{operator-round-round}{\round{\placeholder}}{\placeholder{} rounded towards the nearest integer, as in \glsentrytext{IEEE} 754-2008 (\code{round})}
